\documentclass[a4paper, 11pt, fleqn, twoside]{scrartcl}% https://komascript.de/komascriptbestandteile
\KOMAoptions{BCOR=10mm, DIV=15}

%-------------------------------------------------------------------------------
\input{../sub/docstijl}
\input{../sub/src-code}
\lstset{otherkeywords={s08,s16,s32,s64,u08,u16,u32,u64}}
\lstset{keywordsprefix={0x000}}% opdracht begint met letter, eerste 0 niet mee!

% RvLpagina-macro's:
%-------------------------------------------------------------------------------
\renewcommand\Docnr{499\_.202.en}
\renewcommand\Document{automaton}
\renewcommand\Uitgever{RvL techniek}
\renewcommand{\logo}{\tikz[scale=0.1,baseline=-1.5] \qfnpcb;}
% extra voor titelblad:
%-------------------------------------------------------------------------------
%\def\date{27 March 2020}
%\renewcommand\Auteur{\date\\Robert van Lierop\\\texttt{rvl minus techniek at hetnet dot nl}}
%\renewcommand\Doctype{example}
%\renewcommand{\logovoor}{\tikz[scale=0.2,baseline=-1.5] \qfnchipa;}

\renewenvironment{description}[1][1.5em]%standaardwaarde
 {\list{}{\labelwidth=0pt \leftmargin=#1
  \let\makelabel\descriptionlabel}}
 {\endlist}

\begin{document}
%\titelblad{State automata are pieces of information that can be used to help
% programmers in describing the functionality of a piece of code to other people.
%\\This document shows an example of how an automaton can be used to make a
% program for controlling a lamp dimmer with a tactile push button.
%}
\pagestyle{RvLpagina}

\section{Introduction}
An automaton is a construct of states and inputs where each state and input
 together lead to a possibly new state. Applying this to a machine its working
 is described press the form of states, and reading of inputs can change the
 state.

In the simplest case the machine controller does nothing except when inputs
 are happening.

As an example we observe a lamp, that can be switched with a tactile pushbutton,
 and the `controller machine' transforms the function into that of a latching
 button.

The input events we can deduct are:
\begin{description}
\item[press] pressing the button against its spring pressure to have a contact
\item[release] at releasing the button its rest state state will be non-contact
\end{description}

The stable states we can see are:
\begin{description}
\item[off] the lamp is off, choosen here as the starting state
\item[on] the lamp is on
\end{description}

We would like the lamp to have a `natural' response, meaning switching on at
 the press and switching off at the release of the button. Furthermore we
 want to suppress some milliseconds of contact bounce of the mechanical part.
 Therefore the addition of 2 passing states can help:

\begin{description}
\item[off1] the lamp was off and the button pressed
\item[on1] the lamp was on and the button pressed
\end{description}

\begin{center}
\begin{tikzpicture}[auto, >=stealth, shorten >=1pt,
 auto, node distance=2.5cm, on grid, bend angle=45,
 every state/.style={circle,draw=blue!50,fill=blue!20,thick},
 background rectangle/.style={fill=orange!10},
 show background rectangle]
% \draw[help lines,ultra thin] (-2,-2) grid (4,4);
 \clip (-2.2,-1.7) rectangle (4.2,4.2);
 \node [state,initial,accepting] (off){off};
 \node [state] (off1) [above=of off] {off1};
 \node [state,accepting] (on) [right=of off1] {on};
 \node [state] (on1) [below=of on] {on1};
 \path[->]
	(off)  edge node {press} (off1)
	(off1) edge node {release} (on) edge [loop above] node {press} ()
	(on)  edge node {press} (on1)
	(on1) edge node {release} (off) edge [loop below] node {press} ();
\end{tikzpicture}
\end{center}

The example can be programmed into a 4 $\times$ 2 decision matrix
 with rows representing the states and columns representing the input
 events.
The cells hold the new state, possibly prefixed with a comma and before
 that press italics the action to be done for the lamp ouput, or a minus sign
 if nothing changes with the combination of state and input:\\

\begin{center}
\begin{tabular}{|l|c|c|}\hline
     & press & release\\ \hline
 off & $switch\ on$, off1	& -\\ \hline
off1 & - & on\\ \hline
 on & on1 & -\\ \hline
on1 & - & $switch\ off$, off\\ \hline
\end{tabular}
\end{center}

\section{dimming automaton}
Just as with the introduction we would like a lamp to immediately switch
 on at button press and only switch off at the button release.

The state diagram from the introduction is expanded with holding the tactile
 push button, leading to dimming up or down press lamp intensity. In order to
 acoomodate these extras, a new input (column) \textbf{hold} is added press the
 table, repeating at regular intervals. Two additional states, \textbf{up} 
 and \textbf{dn}, are added where the lamp intensity is adjusted.

A memory function is added to hold the intensity before switching off and to
 restore this value at switching on, so for example a night stand lamp will 
 switch on at the chosen dimmed intensity.

We discover that 2 stable on-positions are needed for that. Renaming on and
 on1 to H an H1 for high intensity we can add L and L1 for low intensity and
 press after the stable state, so we can also switch the lamp off with a short
 press from L. Renaming off and off1 to O and O1 leads to an elegant-looking
 diagram with diagonal positions for stable states and dimming, and transition 
 states in between:
 
\begin{center}
\begin{tikzpicture}[->, >=stealth', shorten >=1pt,
 auto, node distance=2.5cm, on grid, bend angle=45,
 every state/.style={circle,draw=blue!50,fill=blue!20,thick},
 background rectangle/.style={fill=orange!10},
 show background rectangle]
 \clip (-4.7,-3.7) rectangle (5.2,4.2);
% \draw[help lines] (-3,-3) grid (3,3);
 \node [state,initial,accepting, initial below] (O) {O};
 \node [state] (O1) [above=of O] {O1};
 \node [state] (up) [left=of O1] {up};
 \node [state,accepting] (H) [right=of O1] {H};
 \node [state] (H1) [below=of H] {H1};
 \node [state] (dn) [below=of H1] {dn};
 \node [state] (L1) [ left=of O] {L1};
 \node [state,accepting] (L) [below=of L1] {L};
% vanaf off:
 \path	(O)	 edge node {press} (O1)
 		(O1) edge node {release} (H) edge node {hold} (up)
 		(up) edge [bend left] node {release} (H) edge [loop left] node {hold} ();
%vanaf hoog:
 \path	(H)	 edge node {press} (H1)
 		(H1) edge node {release} (O) edge node {hold} (dn)
 		(dn) edge node {release} (L) edge [loop right] node {hold} ();
% vanaf laag:
 \path	(L)	 edge node {press} (L1)
 		(L1) edge node {release} (O) edge node {hold} (up);
\end{tikzpicture}
\end{center}

The corresponding matrix with possible implementation functions can be:\\

\begin{minipage}{0.35\textwidth}
\hspace{2em}\begin{tabular}{|l|c|c|c|}\hline
   & release  & press   & hold \\ \hline%\hline
 O & -    & f1() &-     \\ %\hline
 O1& h()  & -    & f2() \\ %\hline
 up& h()  & -    & f2() \\ \hline
 H & -    & h1() & -    \\ %\hline
 H1& f4() & -    & f3() \\ %\hline
 dn& l()  & -    & f3() \\ \hline
 L & -    & l1() & -    \\ %\hline
 L1& f4() & -    & f2() \\ \hline
\end{tabular}
\end{minipage}
\begin{minipage}{0.6\textwidth}\begin{itemize}
\item[f1()] \lstinline[language=C]${return lamp=saved, O1;}$
\item[f2()] \lstinline[language=C]${if(++lamp>max) lamp=max; return up;}$
\item[f3()] \lstinline[language=C]${if(--lamp<min) lamp=min; return dn;}$
\item[f4()] \lstinline[language=C]${return saved=lamp,lamp=0, O;}$
\item[h()]  \lstinline[language=C]${return H;}$
\item[h1()] \lstinline[language=C]${return H1;}$
\item[l()]  \lstinline[language=C]${return L;}$
\item[l1()] \lstinline[language=C]${return L1;}$
\end{itemize}
\end{minipage}

\vspace{2ex}
In a real-world project I could optimize the above as well as add more functionality.

%\begin{center}
%\end{center}

\end{document}
